On Sample Size Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-09-01 Manuscript/manuscript.tex
Small Update
went over exp simulation; tuned floats

\renewcommand{\textfraction}{0.10}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\floatpagefraction}{0.90}

@@

@@

\jy{What was $B$?} \mc{addressed}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of an AR(1) process with an 
    underlying standard normal distribution with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The 
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:sigma}
\end{figure}

@@

An additional simulation study in which block bootstrap is used to estimate
\jy{not an AR(1) process...} \mc{addressed}
parameters of a marginally transformed AR(1) process with an underlying 
exponential distribution is discussed in the Appendix.

@@

This appendix concerns the performance of block bootstrap when used to estimate
the mean, standard deviation, and the lag-1 autocorrelation coefficient of a 
AR(1) process with an underlying unit exponential marginal distribution. The
motivation for this additional simulation study was to observe if the results of 
the main simulation study will change if the underlying marginal distribution is
not normal.

We generated time series $X_t$ from a 1st order autoregressive 
(AR(1)) process as we did previously, but transformed the data from a standard
normal distribution to an exponential distribution with rate $\lambda = 1$, so 
that for $t \in \{1, \ldots, n\}$, we can define:
\begin{equation*}
W_t = F^{-1}(\Phi(X_t), 1),
\end{equation*}
where $F^{-1}(p, \lambda)$ is the quantile function for the exponential 
distribution with rate parameter $\lambda$.

The simulation design was the same, except the target parameters of $W_t$ were:
1) $\mu = 1$, the mean of $W_t$;
2) $\sigma_w = 1$, the standard deviation of $W_t$; and
3) $\phi$, the lag-1 autocorrelation coefficient of $W_t$.
Let $\rho$ be the lag-1 autocorrelation coefficient of $X_t$.
Although we may have set $\rho$ to these five levels for $X_t$: 
$\{-0.4, -0.2, 0.0, 0.2, 0.4\}$, because of the transformation,
the new five true autocorrelation coefficients for $W_t$ were different.

To calibrate the copula parameters, we first constructed a multivariate
exponential distribution from a normal copula with 
$\rho \in \{-0.4, -0.2, 0.0, 0.2, 0.4\} $using the \texttt{copula} package.
\citep{hofert2018ealements} Then we computed $E[XY]$, which 
requires double integration using the \texttt{pracma} package. \citep{pracma} 
Finally, we obtained true $\phi = E[XY] - 1$, which was used as the target
parameter when evaluating CI performance when estimating $\phi$.
After this step, we again plotted empirical coverage 
rates for $\mu$, $\sigma_w$, and $\phi$, as well as 95\% confidence intervals of 
the real coverage. These are displayed in 
\textbf{Figures~\ref{fig:exp_mu}--\ref{fig:exp_phi}}.

@@

Additionally, another set of 50 CIs was generated for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_pho_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for 
    samples of $n \in \{800, 1600\}$ for a marginally transformed AR(1) process 
    with $\phi = 0.4$
    \jy{not accurate} \mc{is this better}
    with an underlying exponential distribution ($\lambda = 1$). For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

It seems that range of sample sizes needed 
to cover parameters of an exponentially distributed AR(1) process is generally 
higher than that needed to cover parameters of a normally
distributed AR(1) process. Despite this, the trends and patterns discussed 
previously with respect
to the performance of different methods and with different parameters are the 
same.
For instance, Student's $t$ CIs still appear to yield higher coverage rates than 
other 
methods. Performance is still generally better when temporal dependence is 
negative rather than positive. Most importantly, percentile, BC, and BCA CIs
still deteriorate in coverage of $\phi$ as sample size increases. As exhibited 
in Figure~\ref{fig:eri}, percentile and BC CIs still suffer from the same bias
issue, whereas recentered percentile CI still work when estimating $\phi$ 
because the 
original point estimator is unbiased. Overall, the results with an underlying 
marginal 
exponential distribution 
are very similar to the results with an underlying normal distribution from the
main study.

\jy{comment on that the recentered percentile CI for autocorrelation still works
  as the point estimator is unbiased.} \mc{addressed}

>>>>

% control floats
%% \renewcommand\floatpagefraction{.9}
%% \renewcommand\topfraction{.9}
%% \renewcommand\bottomfraction{.9}
%% \renewcommand\textfraction{.2}
%% \setcounter{totalnumber}{50}
%% \setcounter{topnumber}{50}
%% \setcounter{bottomnumber}{50}

@@

\usepackage{physics} % for \dd

@@

@@

\begin{figure}[bp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of an AR(1) process with an 
    underlying standard normal distribution with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The 
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:sigma}
\end{figure}

@@

An additional simulation study in which block bootstrap is used to estimate the
parameters of a stationary series with exponential distribution distribution is
reported in the Appendix. The results are similar.

@@

This appendix concerns the performance of block bootstrap when used to estimate
the mean, standard deviation, and the lag-1 autocorrelation coefficient of a
stationary series with an exponential marginal distribution. The motivation of
for this study was to investigate if the results are robust to nonnormal
marginal distributions.

The series were generated by marginally transforming the
AR(1) series  $X_t$in the simulation study by
\begin{equation*}
W_t = F^{-1}[\Phi(X_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the unit exponential 
distribution. The true mean and standard deviation parameters are~1. The lag-1
autocorrelation coefficient is not invariant to the
transformation\citep{hofert2018ealements}, but its value can be obtained by
\begin{equation*}
\int_\infty^\infty \int_\infty^\infty F^{-1}[\Phi(x)] F^{-1}[\Phi(y)] g_2(x, y; \phi) \dd x \dd y - 1,
\end{equation*}
where $g_2(x, y; \phi)$ is the density of a standard bivariate normal
distribution with correlation~$\phi$. We kept the configuration of
$\phi \in \{-0.4, -0.2, 0.0, 0.2, 0.4\}$.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_phi_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for the
    mean, standard deviation, and lag-1 autocorrelation coefficient
    of a stationary series with marginal exponential distribution
    obtained by transforming an AR(1) process with $\phi = 0.4$ with
    sample size $n \in \{800, 1600\}$. For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

\jy{change the captions of the other three figures accordingly}

The empirical coverage rates for $\mu$, $\sigma_w$, and $\phi$, as well as 95\%
confidence intervals of the real coverage are displayed in 
\textbf{Figures~\ref{fig:exp_mu}--\ref{fig:exp_phi}}. 
Additionally, another set of 50 CIs was generated for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$ in Figure~\ref{fig:eri}.

It appears that a greater sample size is generally required for the bootstrap
CIs to cover the mean and standard deviation parameters in the exponential
margin case than in the normal margin case. However, the trends and patterns
discussed in the main text regarding the performance of various methods and
diverse parameters remain consistent. For example, Student's $t$ confidence
intervals still exhibit higher coverage rates in comparison to alternative
methods. Performance continues to be more favorable when temporal dependence is
negative rather than positive. Of particular significance, the percentile, BC,
and BCA confidence intervals still display a decline in coverage accuracy for
the lag-1 autocorrelation coefficient as sample size increases as demonstrated
in Figure~\ref{fig:eri}. Both the percentile and BC intervals persist in
manifesting the same bias issue. On the other hand, the recentered percentile
confidence interval continues to be effective in estimating $\phi$ due to the
inherent unbiasedness of the original point estimator. In sum, the findings for
series that are marginally exponentially distributed closely mirror those
attained for series that are marginally normally distributed.
